Molecular dynamics simulation of synchronization in driven particles

### 2021-02-25 manuscript/manuscript_ajp_2020.tex
rewrote phase section

%Christiaan Huygens 
%[has been adapted into many useful technologies...]
Synchronized phase-locking or mode-locking 
first appeared in the scientific literature
with 
Huygens' 1665 experiments on
motions of synchronized 
pendula in wall-mounted clocks \cite{Bennett2002}.
%Even when started out of phase, 
%Huygens demonstrated the tendency of
%coupled pendula tend to swing in time
%due to interactions through the wall 
%phase locking
A locked-mode is an integer frequency ratio.
In Huygens clocks,
%even when started out of phase, 
the pendula were observed swing at same rate -- a 1:1 mode.
For a higher modes,
consider simple pendula of different lengths,
a 2:1 mode occurs when a pendulum four
times the length of another swings with twice the period.
%
%to a mode or frequency ratio  
%of an integer number \cite{Bak1986}.
%Dynamical mode locking
%The phenomena

%\section{Colloidal Models}
%\label{sec:colloids}

@@

A single particle oscillator
in a potential well is 
like a skateboarder in a half-pipe or
a child on a swing.
 A confined oscillator may 
 synchronize its location
 to the periodic pattern of the external drive,
 moving back and forth in time with
 the beat,
 or moving between substrate minima.
 %The relationship between the driving force and
 %landscape can change the pattern of motion.  
 A constant or 
 dc drive the landscape modulates 
 the particle velocity.
 Below some threshold  
 the dc force is not strong enough to push the particle
 across a potential maximum so the average velocity is zero,
 a phenomena referred to as pinning \cite{Reichhardt2017}.
 Above the pinning threshold,
 a particle subject to a constant
 drive force will increase its speed at a rate propotional
 to the external drive.  
 %and dissipating energy from the system at all driving forces.
 When the applied force varies periodically,
 the ac drive can 
 cause the particle to hop back and forth across
 the landscape minima.
 Many synchronized patterns occur
 controlled by the substrate period,
 leading to 
 %A sinusoidal or ac drive can create
 mode-locking,
 where the average particle velocity
 is fixed for a range of dc drive forces \cite{Reichhardt2015}.
 %This synchronization phenomena
 %is useful for understanding the complex dynamical 
 %systems.

%Numerical modeling of colloids can provide mechanistic insight
%that can be difficult to achieve in experimental conditions
%where Brownian motion and other sources of noise dominate.
%external forces
%The
%dynamics of particles subject to 
%an applied external force apply to many physical systems.
%For instance the flow of charges in a conductor,
%or cells responding to a chemical gradient.
%The external force 
%increases the diversity of dynamical behaviors,
%and can cause particles to flow in
%a variety of non-linear complex patterns
%that may be 
%synchronized.
%Disordered chaotic dynamics are also possible,
%where irregular, unpredictable time evolution of
%nonlinear systems occurs in mechanical oscillators \cite{chaos}.
%-------------------------------------------------------------

%Given the prevalence of synchronization,
%it is the subject of many computational models.
Here we perform numerical studies 
on the synchronized dynamics
of confined particles driven over
a washboard shaped potential energy landscape.
%Numerical studies %of harmonic systems
%can yield insights
%to the mechanics of naturally oscillating systems.
%We choose this model for its
%relevance to condensed matter systems
%and ease of simulation.
%---------------------------------------------------------------
%kinks and nanofriction
%-----------------------------------------------------------------
%transition from synchronized to chaotic
%describe types of physical systems - particle types
%%In this work,
%we perform 
%numerical simulations of confined, driven particles
%to model a variety of physical phenomena.
%-----------------------------------------------------------------------
%paper outline
%In Section~\ref{sec:colloids} we describe colloid particles,
%an relevant
%experimental analog of our simulations.
We describe
our molecular dynamics model for a single particle in Section~\ref{sec:MD}.
We choose this model for its
relevance to condensed matter systems
and ease of simulation.
%In this section, we restrict our discussion to a single particle,
%and include additional information
%for multiple particles in Section~\ref{sec:sync}.
%including various  and confining environments.
We summarize
our results 
%in Section~\ref{sec:results} 
including synchronized motion of a single confined particle
driven across a periodic landscape in 
Section~\ref{sec:one}.
%In Section~\ref{sec:sync}
%we 
%extend the model to include multiple particles,
%and show how confined 
%interacting particles respond to the same applied driving force.
%including stationary propagation of high density kinks
%in Sec.~\ref{sec:kink}.
%We include code to simulate
%and visualize the landscape and dynamics in this section
%and supplementary material.
%we
%demonstrate how uniform environments and applied forces
%and create synchronized flow patterns.
%We present these results using standard tools of non-linear oscillators
%such phase diagrams of velocity versus position.
%PHASE DIAGRAMS SHOWING COUPLING, OR QUASI-PERIODIC COUPLING
%INDICATE COOPERATIVE MOTION IN THE Q1D SUBSTRATE
%In Section~\ref{sec:quasiperiod},
%we show how an aperiodic landscape modifies the particle dynamics.
%Finally we explore the transition to chaotic dynamics in 
%Section~\ref{sec:chaos}
%and conclude 
We include 
exercises for interested students 
in Section~\ref{sec:problems}.
In Section~\ref{sec:conclusion}
we describe how our results apply 
to physical systems
such as dusty plasmas, superconducting vortices and Josephson junctions.

%Here we model the microscopic dynamics of phase locking in a colloidal model system using molecular dynamics (MD).

\section{Molecular Dynamics Simulation}
\label{sec:MD}

We use a classical model for 
studying the dynamics of $N$ interacting particles,
using the net force on each particle to calculate
its trajectory.
Particles are confined in a two-dimensional (2D) 
simulation of area $A = L \times L$ where $L=36.5 a_0$
where $a_0$ is a dimensionless unit of length.
An individual particle $i$ has
position $\vec{r}_i = x_i \hat{x} + y_i \hat{y}$
and velocity $\vec{v}_i = d\vec{r}_i/dt$.
The edges of the system are treated with
periodic boundary conditions
such that a particle leaving the edge of the system is mapped
back to a position within the simulation boundaries 
by the transformation $x_i+L \rightarrow x_i$ and $y_i+L \rightarrow y_i$.
%and particles interact across PBC.
We show a schematic of the system in Fig.~\ref{fig:landscape0}(a).
The units of the simulated variables are summarized in Table ~\ref{tab:1}.

@@

 Particles are subject an external time-dependent driving force
$\vec{F}^{D}(t)$
applied parallel to the y-direction.
We model this force as
\begin{equation}
  \vec{F}^{D}(t) = [F^{dc} + F^{ac} \sin(\omega t)] \hat{y},
    \label{eq:drive}
\end{equation}
with %parameters including
a constant component $F^{dc}$,
a time dependent component with amplitude $F^{ac}$
and frequency $\omega = 2 \pi f$.
%in the positive $y-$direction.
%but it may 
%be zero or negative depending the the parameters and time.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%RAMP!!!
%%%When $F^{dc}$ is non-zero,  we slowly increase $F^{dc}$ from zero to avoid transient behavior by ensuring the particle is subject to a smoothly varying force. We illustrate the increase of $F^{dc}$ in Fig.~\ref{fig:0}(a).
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%because
The inertia of 
small particles is reduced by interactions
with fluid particles \cite{Purcell1977}.
In our model
we assume 
colloids are overdamped
-- i.e. colloids are suspended in a continuous viscous fluid
that dissipates energy %at a rate
such 
that the particles do not accelerate.
%that greatly impedes their motions.
Energy dissipation from the fluid is modeled
with a frictional force on particle with velocity $\vec{v}_i$ as
%that removes energy from
%the particles,
%which we model as
$\vec{F}^{drag}_i = -\eta \vec{v}_i$
%proportional to the particle velocity 
with a friction coefficient $\eta$
proportional to the fluid viscosity.
We discuss friction models for
spheres moving through fluids in 
%Section~\ref{sec:problems}
Exercise~\ref{ex:reynolds}. %for more information
%We model the particle dynamics  with an overdamped equation of motion. Models with inertia set to zero are appropriate for moving through a high viscosity fluid
\begin{center}
\begin{figure}[h!]
\centering
\includegraphics[width=\columnwidth]{landscape.pdf}
\caption{
  Schematic of the simulation
  of a single particle
  driven across a washboard potential
  energy landscape.
  The period of the landscape is $\lambda = L/N_p$,  
  %is marked in red,
  where $N_p = 3$.  
  The force due to the landscape %$\vec{F}^l$
  is calculated from the gradient of the potential
  energy 
  $\vec{F}^l = -\nabla U(\vec{r})$.
  (a) View of the $x-y$ plane. %of the simulation
  The time-dependent applied driving force $\vec{F}^D$
  is parallel to the y-axis.
  The landscape is shown with a contour plot,
  with maxima in the potential energy marked in red
  and minima marked in blue.
  A particle is shown in the region between minima and maxima
  subject to competing forces of the landscape and applied driving force.
  (b) The potential energy function
  along the y-axis $U_l(y)$.  
  The particle in (a) is shown at the same $y-$position.
  The slope of $U_l(y)$ is the 
  magnitude of force $\vec{F}^l$. %is calculated
  %as the slope of this function,
  %directed along the $y-$direction in panel (a).
  }
\label{fig:landscape0}
\end{figure}
\end{center}
Newton's second law for an individual particle
%$m \vec{a}_i = \sum \vec{F}$
is simplified
by the assumption $\vec{a}_i$ is zero. % in the overdamped model.
%Newton's second law can be rearranged %for the velocity
%to
The overdamped equation of motion for an isolated particle is
%solved for velocity 
\begin{equation}
  \eta \vec{v}_i = \vec{F}^l_{i}(\vec{r}_i) + \vec{F}^{D}(t).
    \label{eq:motion}
\end{equation}
with  $\eta = 1$ in units of $v_0 / F_0$. 
%(see Table~\ref{tab:1}).
%and Exercise~\ref{ex:reynolds}.
%In %Section~\ref{sec:problems}
%Exercise~\ref{ex:n2l}
%we invite the reader to confirm this result.
The equation of motion provides a direct calculation of the velocity
of an individual particle from its location $\vec{r}_i$ %in the simulation
and the simulation time.

The molecular dynamics simulation is controlled by a $for()$ loop
which runs from an initial to maximum integer number of time steps.
%$t_i$.
Each integer time step interval %$t_{i+1}-t_i$
represents a simulation time of $\Delta t$ 
simulation units $\tau$.  %as described in Table 1
At each time step
we evaluate the net force on each particle as a function of its position
$\vec{r}_i(t)$
and then integrate
the equation of motion to move particles
to an updated position.
%$\vec{r}_i(t+\Delta t)$.%
Since the acceleration as zero,
the integration of the equation of motion
is performed via 
the Euler method 
\begin{equation}
  \vec{r}_i(t+\Delta t) = \vec{v}_i(t) \Delta t + \vec{r}_i(t)
    \label{eq:euler}
\end{equation}
for a small time step $\Delta t = 0.01 \tau$.
In %Section~\ref{sec:problems}
Exercise~\ref{ex:euler}
we describe 
the numerical methods for 
solving differential equations.
%

\begin{table}[h!]
\centering
\caption{Simulation parameters and units with comparable
  experimental values~\cite{Juniper2015}.}
  %Blanks indicate you are a moron
  %units, rather than parameters that
  %are not easily represented by a single value.}
%{\it might be more appropriate to match Juniper parameters more precisely given our focus...}}
\begin{ruledtabular}
\begin{tabular}{c c c } %p{5cm}}
Quantity & Simulation Unit & Experimental value\\
\hline
length &  $a_0 = 1$ & $ a_0 \sim 1.5 \mu m$\\
energy & $E_0 = 1$ & \\ %$ < E_0 <$ \\
%electric potential & $V(r_{ij}) = E_0/r_{ij}$ \\
%energy & $E_0 = q^2{Z^*}^2/4\pi \epsilon \epsilon_0 a_0$ \\
%dimensionless interaction strength & $q$ \\
%effective colloidal charge & $Z^*$ \\
%solvent dielectric constant & $\epsilon \epsilon_0$\\
force & $F_0 = E_0 / a_0$ & \\ %$<F_0<$\\
time &  $ \tau = \eta a_0 / F_0 $ \\ %& $ < \tau <$\\
velocity &  $ a_0 / \tau $ &  $v \sim 5 \mu m/s $\\
driving period & $T = 100 \tau$ & $T \sim 1.3$ s \\ %frequency 0.75 Hz
viscosity coefficient & $\eta = F_0 \tau / a_0 $ & $\eta \sim 10^{-3}$ Pa-s \\
%juniper uses water–ethanol mixture
substrate period & $\lambda = 1.825 a_0$ & $\lambda = 3.5 \mu m$ \\
substrate amplitude & $U_0 = 0.058 E_0$ & $U_0 = 25 k_B T \sim 2 J$ \\
%they have 75 k_B T as the bottom of -U_0
%A_p = 0.1 = N_p pi U_0/L.  this is how calculated U_0
temperature & $T_0 = 0$ \footnote{to learn more about the effects
  of temperature on this simulation
  see Ex.~\ref{ex:brownian} where we study the limit
  $U_0 \sim k_B T$.} & $T \sim 300K $ \\

%FROM JUNIPER2015
%l¼3.5 mm and a laser power per trap of WE0.75mW (trap stiffness: k=3.8\times10^7 kg /s^2; trap depth: V0 = 75 k_B T
%ref51
%Juniper, M. P. N., Besseling, R., Aarts, D. G. A. L. & Dullens, R. P. A. Acousto- optically generated potential energy landscapes: potential mapping using colloids under flow. Opt. Express 27, 28707–28716 (2012).
%U(x) measured in ratio of k_BT from -75 to -25
%0 < vAC < 8 micrometer/s and frequency f=0.25 to 0.75 Hz.
%0 < vDC < 5 micrometer/s
%convert to force units?
%check when steps exist - in Juniper - steps don't exist for n=1or n=3, even though steps exist for n¼0,2,4,5 and 6.
%flexible chains
%dipole interaction $\Gamma \propto B^2$
%\Gamma =15 (B=0.43mT) and a stiff chain at \Gamma=392 (B=2.2mT).
%Kink
%16 particles in a strong potential energy landscape of 15 minima (trap spacing: l¼5.5 mm; laser power per trap: WE1.75 mW; trap stiffness: k¼8.7?10?7 kg s?2; trap depth: V0¼200 kBT),
%JUNIPER2017
%The driving force is provided by a PI-542.2CD piezo-stage, controlled using the LabView interface, at driving 0.4 ≤ FAC/ζ ≤ 14 µm s−1, and a frequency of 1 velocities of 0.05 ≤ FDC/ζ ≤ 8 µm s−1. AC driving veloc
%Volpe AJP Parameters
%We will consider a silica microparticle in water with radius R=1micro m, mass m=11picograms, viscosity eta=0.001 Ns/m^2; $\gamma=6\pi \eta R$,  T=300K, and \tau=0.6\micro s.
%thermofisher off the shelf paramagnetic beads
%https://www.thermofisher.com/us/en/home/references/protocols/proteins-expression-isolation-and-analysis/protein-isolation-protocol/m-270-carboxylic-acid.html
%~2\times10^9 beads per 30 mg -> 1.5 \times 10^-11 g, 15 ng

\end{tabular}
\end{ruledtabular}
\label{tab:1}
\end{table}

%\section{Results}
%\label{sec:results}
% We demonstrate how a single particle (Sec.~\ref{sec:one})
% and many particles
% move (Sec.~\ref{sec:sync})
% in response to this applied force in a variety of environments.
% In Sec.~\ref{sec:kink} we set $F^{dc}$
% to zero and track the motion
% of a high density area of a particle chain
% (i.e. kink dynamics).
 %The driving force does add energy into the system, and some of it is lost.
%
\section{Mode-locking of a single particle}
% In subsection headings, only the first word is capitalized.
\label{sec:one}

Here we drive 
a single particle %on this landscape
%responds to the applied driving force
%by moving
across the landscape %maxima
to characterize its motion.
%synchronized with 
%the period of $F^D(t)$.
%applied driving force.
%along the y-direction
%({\it redundant, just refer back to eq. in Model?})
%and $V_{0y}=0.2$.
%This is illustrated in Fig.~\ref{fig:1}(a) where
%the red (blue) regions show local maxima (minima).
%
The numerical implementation of the landscape 
is calculated with Eq.~\ref{eq:ysubstrate} as 
\begin{equation}
  \label{eq:force}
  F^l_y(y) = -A_{p} \sin{(2 \pi y / \lambda)} 
\end{equation}
where the force is scaled with parameter $A_{p} = 2\pi U_0/\lambda$.
In this section we fix the landscape parameters
to $A_{p} = 0.1 F_0$ %force units
with $N_p=20$ minima %in the landscape
corresponding to a spatial period $\lambda = 1.825 a_0$.
The competition between the driving force and landscape potential
can produce a variety of hopping patterns in the particle motion. 
%In Fig.~\ref{fig:0}
%reaching a minimum position
%of $y/\lambda = 1$
%over the total period $T = 100$.
%In Fig. ~\ref{fig:0}
%the average displacement during one time period is a single wavelength
%of the substrate $y(t_0+T) - y(t_0) = \lambda$.
%single dc case with figure
%dc_current 0.01
%ac_current 0.05
%ac_frequency 0.01
%fp_y 0.1
%
The relative values of $F^{ac}$, $F^{dc}$ and $A_p$
control the rate and distance a  particle moves 
forward and backward in the landscape.
[edit and move:] The addition of an oscillating force
can lead to interference effects
that create stepped force-velocity curves.
Useful model - essentially an overdamped driven pendulum -
under certain parameters.
The steps occur
due to the phase locking between lattice substrate
and applied driving force.
Step widths depend on AC drive.
Discrete modes.
%Shapiro steps wash out to sloped regions at
%high vortex numbers,
%so what is the analog in my system?
%need to sweep through more parameters to find and report the flat regions
%not fully phase locked.

>>>>

%Christiaan Huygens 
%[has been adapted into many useful technologies...]
Synchronized phase-locking or mode-locking 
first appeared in the scientific literature
with 
Huygens' 1665 experiments on
motions of synchronized 
pendula in wall-mounted clocks \cite{Bennett2002}.
%Even when started out of phase, 
%Huygens demonstrated the tendency of
%coupled pendula tend to swing in time
%due to interactions through the wall 
%phase locking
A locked-mode is an integer frequency ratio.
In Huygens clocks,
%even when started out of phase, 
the pendula were observed swing at same rate -- a 1:1 mode.
For a higher modes,
consider simple pendula of different lengths,
a 2:1 mode occurs when a pendulum four
times the length of another swings with twice the period.
%
%to a mode or frequency ratio  
%of an integer number \cite{Bak1986}.
%Dynamical mode locking
%The phenomena
%\section{Colloidal Models}
%\label{sec:colloids}

@@

A single particle oscillator
in a potential well is 
like a skateboarder in a half-pipe or
a child on a swing.
 A confined oscillator may 
 synchronize its location
 to the periodic pattern of the external drive,
 moving back and forth in time with
 the beat,
 or moving between substrate minima.
 %The relationship between the driving force and
 %landscape can change the pattern of motion.  
 A constant or 
 dc drive the landscape modulates 
 the particle velocity.
 Below some threshold  
 the dc force is not strong enough to push the particle
 across a potential maximum so the average velocity is zero,
 a phenomena referred to as pinning \cite{Reichhardt2017}.
 Above the pinning threshold,
 a particle subject to a constant
 drive force will increase its speed at a rate propotional
 to the external drive.  
 %and dissipating energy from the system at all driving forces.
 When the applied force varies periodically,
 the ac drive can 
 cause the particle to hop back and forth across
 the landscape minima.
 Many synchronized patterns occur
 controlled by the substrate period,
 leading to 
 %A sinusoidal or ac drive can create
 mode-locking,
 where the average particle velocity
 is fixed for a range of dc drive forces \cite{Reichhardt2015}.
 %This synchronization phenomena
 %is useful for understanding the complex dynamical 
 %systems.
 The model presented in this paper
 resembles an overdamped driven pendulum. 
%Numerical modeling of colloids can provide mechanistic insight
%that can be difficult to achieve in experimental conditions
%where Brownian motion and other sources of noise dominate.
%external forces
%The
%dynamics of particles subject to 
%an applied external force apply to many physical systems.
%For instance the flow of charges in a conductor,
%or cells responding to a chemical gradient.
%The external force 
%increases the diversity of dynamical behaviors,
%and can cause particles to flow in
%a variety of non-linear complex patterns
%that may be 
%synchronized.
%Disordered chaotic dynamics are also possible,
%where irregular, unpredictable time evolution of
%nonlinear systems occurs in mechanical oscillators \cite{chaos}.
%-------------------------------------------------------------

%Given the prevalence of synchronization,
%it is the subject of many computational models.
Here we perform numerical studies 
on the synchronized dynamics
of confined particles driven over
a washboard shaped potential energy landscape.
%Numerical studies %of harmonic systems
%can yield insights
%to the mechanics of naturally oscillating systems.
%We choose this model for its
%relevance to condensed matter systems
%and ease of simulation.
%---------------------------------------------------------------
%kinks and nanofriction
%-----------------------------------------------------------------
%transition from synchronized to chaotic
%describe types of physical systems - particle types
%%In this work,
%we perform 
%numerical simulations of confined, driven particles
%to model a variety of physical phenomena.
%-----------------------------------------------------------------------
%paper outline
%In Section~\ref{sec:colloids} we describe colloid particles,
%an relevant
%experimental analog of our simulations.
We describe
our molecular dynamics model for a single particle in Section~\ref{sec:MD},
The model
closely resembles
a damped driven pendulum,
and is easy to simulate yet relevant
to condensed matter systems.
%In this section, we restrict our discussion to a single particle,
%and include additional information
%for multiple particles in Section~\ref{sec:sync}.
%including various  and confining environments.
We summarize
our results 
%in Section~\ref{sec:results} 
including synchronized motion of a single confined particle
driven across a periodic landscape in 
Section~\ref{sec:one}.
%In Section~\ref{sec:sync}
%we 
%extend the model to include multiple particles,
%and show how confined 
%interacting particles respond to the same applied driving force.
%including stationary propagation of high density kinks
%in Sec.~\ref{sec:kink}.
%We include code to simulate
%and visualize the landscape and dynamics in this section
%and supplementary material.
%we
%demonstrate how uniform environments and applied forces
%and create synchronized flow patterns.
%We present these results using standard tools of non-linear oscillators
%such phase diagrams of velocity versus position.
%PHASE DIAGRAMS SHOWING COUPLING, OR QUASI-PERIODIC COUPLING
%INDICATE COOPERATIVE MOTION IN THE Q1D SUBSTRATE
%In Section~\ref{sec:quasiperiod},
%we show how an aperiodic landscape modifies the particle dynamics.
%Finally we explore the transition to chaotic dynamics in 
%Section~\ref{sec:chaos}
%and conclude 
We include 
exercises for interested students 
in Section~\ref{sec:problems}.
In Section~\ref{sec:conclusion}
we describe how our results apply 
to physical systems
such as dusty plasmas, superconducting vortices and Josephson junctions.
%Here we model the microscopic dynamics of phase locking in a colloidal model system using molecular dynamics (MD).

\section{Molecular Dynamics Simulation}
\label{sec:MD}
We use a classical model for 
studying the dynamics of $N$ interacting particles,
using the net force on each particle to calculate
its trajectory.
Particles are confined in a two-dimensional (2D) 
simulation of area $A = L \times L$ where $L=46.\bar{6} a_0$
where $a_0$ is a dimensionless unit of length.
An individual particle $i$ has
position $\vec{r}_i = x_i \hat{x} + y_i \hat{y}$
and velocity $\vec{v}_i = d\vec{r}_i/dt$.
The edges of the system are treated with
periodic boundary conditions
such that a particle leaving the edge of the system is mapped
back to a position within the simulation boundaries 
by the transformation $x_i+L \rightarrow x_i$ and $y_i+L \rightarrow y_i$.
%and particles interact across PBC.
We show a schematic of the system in Fig.~\ref{fig:landscape0}(a).
The units of the simulated variables are summarized in Table ~\ref{tab:1}.

@@

 Particles are subject an external time-dependent driving force
$\vec{F}^{d}(t)$
applied parallel to the y-direction.
We model this force as
\begin{equation}
  \vec{F}^{d}(t) = [F^{dc} + F^{ac} \sin(\omega t)] \hat{y},
    \label{eq:drive}
\end{equation}
with %parameters including
a constant component $F^{dc}$,
a time dependent component with amplitude $F^{ac}$
and frequency $\omega = 2 \pi f$.
%in the positive $y-$direction.
%but it may 
%be zero or negative depending the the parameters and time.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%RAMP!!!
%%%When $F^{dc}$ is non-zero,  we slowly increase $F^{dc}$ from zero to avoid transient behavior by ensuring the particle is subject to a smoothly varying force. We illustrate the increase of $F^{dc}$ in Fig.~\ref{fig:0}(a).
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%because
The inertia of 
small particles is reduced by interactions
with fluid particles \cite{Purcell1977}.
%In our model
We assume 
colloids are overdamped
-- suspended in a continuous viscous fluid
that dissipates energy %at a rate
so the particles do not accelerate.
Newton's second law for an individual particle
%$m \vec{a}_i = \sum \vec{F}$
is simplified
by the assumption $\vec{a}_i$ is zero. % in the overdamped model.
%Newton's second law can be rearranged %for the velocity
%to
The overdamped equation of motion for
the velocity $\vec{v}_i$ of 
an isolated particle is
%solved for velocity 
\begin{equation}
  \eta \vec{v}_i = \vec{F}^l_{i}(\vec{r}_i) + \vec{F}^{d}(t).
    \label{eq:motion}
\end{equation}
with friction coefficient $\eta = 1$ in units of $v_0 / F_0$. 
The term $-\eta \vec{v}_i$
is a drag force used model 
energy dissipation from the fluid. 
%that greatly impedes their motions.
%proportional to the particle velocity 
%with a $\eta$
%proportional to the fluid viscosity.
We discuss friction models for
spheres moving through fluids in 
%Section~\ref{sec:problems}
Exercise~\ref{ex:reynolds}. %for more information
%We model the particle dynamics  with an overdamped equation of motion. Models with inertia set to zero are appropriate for moving through a high viscosity fluid

The equation of motion provides a direct calculation of the velocity
of an individual particle from location $\vec{r}_i$ %in the simulation
and the simulation time.
The molecular dynamics simulation is controlled by a $for()$ loop
which runs from an initial to maximum integer number of time steps.
%$t_i$.
%Each integer time step interval %$t_{i+1}-t_i$
%represents a simulation time of $\Delta t$.
%simulation units $\tau$.  %as described in Table 1
At each time step
we evaluate the net force on each particle as a function of its position
$\vec{r}_i(t)$
and then integrate
the equation of motion to move particles
to an updated position.
%$\vec{r}_i(t+\Delta t)$.%
Since the acceleration as zero,
the integration of the equation of motion
is performed via 
the Euler method 
\begin{equation}
  \vec{r}_i(t+\Delta t) = \vec{v}_i(t) \Delta t + \vec{r}_i(t)
    \label{eq:euler}
\end{equation}
for a time step $\Delta t = 0.1 \tau$.
In %Section~\ref{sec:problems}
Exercise~\ref{ex:euler}
we describe 
the numerical methods for 
solving differential equations.
%
%------------------------------------------------------------------
\begin{figure} % [h!]
\centering
\includegraphics[width=\columnwidth]{landscape.pdf}
\caption{
  Schematic of the simulation of a single particle
  driven across a washboard potential energy landscape.
  %The period of the landscape is $\lambda = L/N_p$,  
  %is marked in red,
  %where $N_p = 3$.  
  %The force due to the landscape %$\vec{F}^l$
  %is calculated from the gradient of the potential energy 
  %$\vec{F}^l = -\nabla U(\vec{r})$.
  (a) View of the $x-y$ plane. %of the simulation
  The time-dependent applied driving force $\vec{F}^d$
  is parallel to the y-axis.
  The landscape is shown with %a contour plot,
  %with
  maxima in the potential energy marked in red
  and minima marked in blue.
  A particle is %in the region between minima and maxima
  subject to competing forces of the landscape and applied driving force.
  (b) The potential energy function
  along the y-axis $U_l(y)$.  
  The particle in (a) is shown at the same $y-$position.
  The slope of $U_l(y)$ is the 
  magnitude of force $\vec{F}^l$. 
  }
\label{fig:landscape0}
\end{figure}
%\section{Results}
%\label{sec:results}
% We demonstrate how a single particle (Sec.~\ref{sec:one})
% and many particles
% move (Sec.~\ref{sec:sync})
% in response to this applied force in a variety of environments.
% In Sec.~\ref{sec:kink} we set $F^{dc}$
% to zero and track the motion
% of a high density area of a particle chain
% (i.e. kink dynamics).
 %The driving force does add energy into the system, and some of it is lost.
%
\section{Mode-locking of a single particle}
% In subsection headings, only the first word is capitalized.
\label{sec:one}

Here we drive 
a single particle %on this landscape
%responds to the applied driving force
%by moving
across the landscape. %maxima
%to explore the qualitative behavior
%to characterize its motion.
%synchronized with 
%the period of $F^d(t)$.
%applied driving force.
%along the y-direction
%({\it redundant, just refer back to eq. in Model?})
%and $V_{0y}=0.2$.
%This is illustrated in Fig.~\ref{fig:1}(a) where
%the red (blue) regions show local maxima (minima).
%
The numerical implementation of the landscape 
is calculated with Eq.~\ref{eq:ysubstrate} as 
\begin{equation}
  \label{eq:force}
  F^l_y(y) = -A_{p} \sin{(2 \pi y / \lambda)} 
\end{equation}
where the force is scaled with parameter $A_{p} = 2\pi U_0/\lambda$.
In this section we fix the landscape parameters
to $A_{p} = 0.1 F_0$ %force units
with $N_p=20$ minima %in the landscape
corresponding to a spatial period $\lambda = 2.3 a_0$.
The competition between the driving force and landscape potential
can produce a variety of hopping patterns in the particle motion. 
%In Fig.~\ref{fig:0}
%reaching a minimum position
%of $y/\lambda = 1$
%over the total period $T = 100$.
%In Fig. ~\ref{fig:0}
%the average displacement during one time period is a single wavelength
%of the substrate $y(t_0+T) - y(t_0) = \lambda$.
%single dc case with figure
%dc_current 0.01
%ac_current 0.05
%ac_frequency 0.01
%fp_y 0.1
%
The relative values of $F^{ac}$, $F^{dc}$ and $A_p$
control the rate and distance a  particle moves 
forward and backward in the landscape.
%Here we explore one case in the low frequency regime,
When $F^d(t) > A_p$, a particle can 
overcome the barrier height of the landscape,
and 
the particle hops between minima in the energy landscape.
When the driving frequency is low,
as in Fig.~\ref{fig:0},  
the driven particle %adjusts
moves %across %its motion to traverse
%the periodic landscape
in a pattern 
with the same frequency 
%with the same rhythm as 
as the time-dependent force $F^d(t)$,
yet is modulated by the landscape period.
We explore changes in frequency and $F^{ac}$ 
in Ex.~\ref{ex:parameters}.
Here we vary $F^{dc}$ 
while holding the remaining parameters fixed.
%------------------------------------------------------------------------
\begin{table} %[h!]
\centering
\caption{Simulation parameters and units with comparable
  experimental values~\cite{Juniper2015,Juniper2017}.}
  %Blanks indicate you are a moron
  %units, rather than parameters that
  %are not easily represented by a single value.}
%{\it might be more appropriate to match Juniper parameters more precisely given our focus...}}
\begin{ruledtabular}
\begin{tabular}{c c c } %p{5cm}}
Quantity & Simulation Unit & Experimental value\\
\hline
length &  $a_0 = 1$ & $ a_0 \sim 1.5 \mu m$\\
energy & $E_0 = 1$ & \\ %$ < E_0 <$ \\
%electric potential & $V(r_{ij}) = E_0/r_{ij}$ \\
%energy & $E_0 = q^2{Z^*}^2/4\pi \epsilon \epsilon_0 a_0$ \\
%dimensionless interaction strength & $q$ \\
%effective colloidal charge & $Z^*$ \\
%solvent dielectric constant & $\epsilon \epsilon_0$\\
force & $F_0 = E_0 / a_0$ & \\ %$<F_0<$\\
time &  $ \tau = \eta a_0 / F_0 $ & $ \tau \sim 3 sec$\\
velocity &  $ a_0 / \tau $ &  $v \sim 5 \mu m/s $\\
%driving period & $T = 100 \tau$ & $T \sim 1.3$ s \\ %frequency 0.75 Hz
%juniper uses water–ethanol mixture
substrate period & $\lambda = 2.3 a_0$ & $\lambda = 3.5 \mu m$ \\
substrate amplitude\footnote{Our substrate is scaled by our force units, while an experimental landscape is scaled by the Brownian motion of particles.} & $A_p = 0.1 F_0$ & $U_0 = 25 k_B T \sim 1 J$ \\ %$U_0 = 0.037 E_0$ 
%they have 75 k_B T as the bottom of -U_0
%A_p = 0.1 = N_p pi U_0/L.  this is how calculated U_0
temperature\footnote{To learn more about the effects
  of temperature on this simulation see Ex.~\ref{ex:brownian} where we study 
  $U_0 \sim k_B T$.} & $T_0 = 0$  & $T \sim 290K $ \\
%FROM JUNIPER2015
%l¼3.5 mm and a laser power per trap of WE0.75mW (trap stiffness: k=3.8\times10^7 kg /s^2; trap depth: V0 = 75 k_B T
%ref51
%Juniper, M. P. N., Besseling, R., Aarts, D. G. A. L. & Dullens, R. P. A. Acousto- optically generated potential energy landscapes: potential mapping using colloids under flow. Opt. Express 27, 28707–28716 (2012).
%U(x) measured in ratio of k_BT from -75 to -25
%0 < vAC < 8 micrometer/s and frequency f=0.25 to 0.75 Hz.
%0 < vDC < 5 micrometer/s
%convert to force units?
%check when steps exist - in Juniper - steps don't exist for n=1or n=3, even though steps exist for n¼0,2,4,5 and 6.
%flexible chains
%dipole interaction $\Gamma \propto B^2$
%\Gamma =15 (B=0.43mT) and a stiff chain at \Gamma=392 (B=2.2mT).
%Kink
%16 particles in a strong potential energy landscape of 15 minima (trap spacing: l¼5.5 mm; laser power per trap: WE1.75 mW; trap stiffness: k¼8.7?10?7 kg s?2; trap depth: V0¼200 kBT),
%JUNIPER2017
%The driving force is provided by a PI-542.2CD piezo-stage, controlled using the LabView interface, at driving 0.4 ≤ FAC/ζ ≤ 14 µm s−1, and a frequency of 1 velocities of 0.05 ≤ FDC/ζ ≤ 8 µm s−1. AC driving veloc
%Volpe AJP Parameters
%We will consider a silica microparticle in water with radius R=1micro m, mass m=11picograms, viscosity eta=0.001 Ns/m^2; $\gamma=6\pi \eta R$,  T=300K, and \tau=0.6\micro s.
%thermofisher off the shelf paramagnetic beads
%https://www.thermofisher.com/us/en/home/references/protocols/proteins-expression-isolation-and-analysis/protein-isolation-protocol/m-270-carboxylic-acid.html
%~2\times10^9 beads per 30 mg -> 1.5 \times 10^-11 g, 15 ng
\end{tabular}
\end{ruledtabular}
\label{tab:1}
\end{table}
